---
title: "Module_3_P_card_setup"
author: "George Batta"
date: "3/3/2021"
output: html_document
---

# Loading and setup
This loads the 6 years of data and stores them in a list object called `data`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(lubridate)
library(stringr)
library(htmlwidgets)


# create a list from these files
# This lists out the names of files with .csv at the end
filenames<-list.files(path="data",pattern=".csv$")


# create an empty list that will serve as a container to receive the incoming files
data<-list()

# create a loop to read in your data
for (i in 1:length(filenames)) {
  data[[i]]<-as_tibble(read_csv(paste(getwd(),"/data/",filenames[i],sep="")))
}


# add the names of your data to the list
names(data)<-filenames
```


This code chunk prints out the column names in each year of data.

```{r Column names}
i=2010
#Column heading cleanup
for (n in data) {
  print(paste("DATA FROM",i,"BELOW"))
  print(colnames(n))
  i=i+1
}



for (i in 1:length(data)) {
  colnames(data[[i]]) <- c("Agency_Number","Agency_Name","Cardholder_Last_Name","Cardholder_First_Initial","Description","Amount","Vendor","Transaction_Date","Posted_Date","MCC")
}

```

Let's now set up our date format tagging and conversion functions
```{r Set up conversion and tagging}
date_case<-function(x) case_when(
      	str_detect(x, "-")==TRUE ~ "A",
  		  str_detect(x, "/")==TRUE  &
		    !str_detect(x,"00$")==TRUE ~  "B",
  	  	str_detect(x, "\\.")==TRUE ~ "C",
  	  	str_detect(x, "/")==TRUE & str_detect(x,":")  ~ "D"
)

convert_date<-function(x) case_when(
  str_detect(x,"-")==TRUE ~ dmy(x),
  str_detect(x, "/")==TRUE  &
	!str_detect(x,"00$")==TRUE ~  mdy(x),
  str_detect(x, "\\.")==TRUE ~ as_date(dmy_hms(x)),
  str_detect(x, "/")==TRUE & str_detect(x,":")  ~ as_date(mdy_hm(x))
)


```

Now, execute!
```{r Execute conversion and Tagging}
for (i in 1:length(data)) {
  data[[i]] <- data[[i]] %>% mutate(across(c(Transaction_Date,Posted_Date),
                                      list(dc=date_case,conv=convert_date)))
}

```

Check frequency of cases among all datasets
```{r Frequency of date cases}

date_case_combined<-data[[1]] %>% 
  select(Transaction_Date_dc,Transaction_Date) 

for (i in 2:length(data)) {
  temp<-data[[i]] %>% select(Transaction_Date_dc,Transaction_Date)  
  date_case_combined<-bind_rows(date_case_combined,temp)
}
date_case_combined<-date_case_combined %>% arrange(Transaction_Date_dc)

date_case_combined %>% 
  group_by(Transaction_Date_dc) %>% 
  summarize(n=n()) %>% 
  mutate(prop=n/sum(n))

```

```{r Merge years}

data[[5]]<-data[[5]] %>% mutate(Amount=as.character(Amount))

p_card_combined<-bind_rows(data)
```


Below, we'll use the `gsub()` command, which has 3 arguments. The syntax is: 

gsub(Regex for finding a pattern or specific characters,
  What you want to substitute in, 
  The column or object you want to perform the substitution on)

We'll then convert Amount to a number, and then dump our intermediate variables...`Transaction Date` and `Posted Date` are the original, character versions of the date variables.

```{r Final cleanup}

p_card_combined <-p_card_combined %>% 
  mutate(Amount=gsub("[$,)]", "", Amount)) %>% 
  mutate(Amount=gsub("\\(", "-", Amount)) %>%
  mutate(Amount=as.numeric(Amount)) %>%
  select(-c(Transaction_Date,Posted_Date,
            Transaction_Date_dc,Posted_Date_dc)) %>% 
  rename(Transaction_Date=Transaction_Date_conv) %>% 
  rename(Posted_Date=Posted_Date_conv) 


#The following statement should be true
sum(is.na(p_card_combined$Transaction_Date))==sum(is.na(date_case_combined$Transaction_Date))

#Dump all but p_card_combined
remove(list=c("data","n","temp","i","filenames","date_case_combined"))


osu14<-p_card_combined %>% 
  filter(Agency_Name=="OKLAHOMA STATE UNIVERSITY") %>% 
  drop_na(Transaction_Date) %>% 
  filter(year(Transaction_Date)==2014) %>% 
  select(-c(Agency_Name,Agency_Number))

```



```{r Benford}
#Benford examples:

osu14 %>% filter(Amount<100 & Amount>=0) %>% ggplot()+
  geom_histogram(mapping=aes(x=Amount))+geom_vline(xintercept=20)+geom_vline(xintercept=10)

osu14 %>% filter(Amount<1000 & Amount>=100) %>% ggplot()+
  geom_histogram(mapping=aes(x=Amount))+geom_vline(xintercept=200)+geom_vline(xintercept=100)

osu14 %>% filter(Amount<10000 & Amount>=1000) %>% ggplot()+
  geom_histogram(mapping=aes(x=Amount))+geom_vline(xintercept=2000)+geom_vline(xintercept=1000)

```


```{r IC1}
#IC 1
osu14 %>% group_by(Cardholder_First_Initial,Cardholder_Last_Name) %>% 
  summarize(Amount=sum(Amount)) %>% filter(Amount>50000) %>% arrange(-Amount)
```


```{r IC2}
#IC 2
osu14 %>% mutate(month=month(Transaction_Date)) %>% group_by(Cardholder_First_Initial,Cardholder_Last_Name,month) %>% 
  summarize(Amount=sum(Amount)) %>% filter(Amount>10000) %>% arrange(month,-Amount)
```


```{r IC3}
#IC 3
osu14 %>% group_by(Cardholder_First_Initial,Cardholder_Last_Name,Vendor,Transaction_Date) %>% 
  mutate(Daily_Amount=sum(Amount)) %>% mutate(count=n()) %>%  filter(count>1 & Daily_Amount>5000) %>% 
  arrange(Transaction_Date,Cardholder_First_Initial,Cardholder_Last_Name,Vendor)

```

```{r Benford Analysis Package}
pacman::p_load(benford.analysis)
benford(osu14$Amount,number.of.digits=1,sign="both",discrete=TRUE,round=3)
```

```{r HWQ1}
osu14$Amount <- as.character(osu14$Amount)

osu14filter <- osu14 %>% filter(Amount > 1.0 | Amount < -1.0)

Num1 <- str_count(osu14filter$Amount,"^1")
count1 <- length(which(Num1 == 1))
count1

Num2 <- str_count(osu14filter$Amount,"^2") 
count2 <- length(which(Num2 == 1))
count2

Num3 <- str_count(osu14filter$Amount,"^3")
count3 <- length(which(Num3 == 1))
count3

Num4 <- str_count(osu14filter$Amount,"^4")
count4 <- length(which(Num4 == 1))
count4

Num5 <- str_count(osu14filter$Amount,"^5")
count5 <- length(which(Num5 == 1))
count5

Num6 <- str_count(osu14filter$Amount,"^6") 
count6 <- length(which(Num6 == 1))
count6

Num7 <- str_count(osu14filter$Amount,"^7")
count7 <- length(which(Num7 == 1))
count7

Num8 <- str_count(osu14filter$Amount,"^8")
count8 <- length(which(Num8 == 1))
count8

Num9 <- str_count(osu14filter$Amount,"^9")
count9 <- length(which(Num9 == 1))
count9

FirstCol <- c(count1,count2,count3,count4,count5,count6,count7,count8,count9)
SecondCol <- c(1,2,3,4,5,6,7,8,9)
Numdf <- data.frame(SecondCol,FirstCol) %>% mutate(Sum=sum(FirstCol)) %>%
  mutate(prop=(FirstCol/Sum)*100) %>%
  mutate(across(prop,round,3)) %>%
  rename("Leading Digit"="SecondCol") %>%
  rename("Frequency"="FirstCol") %>%
  rename("Proportion %"="prop")
Numdf
```
It looks like there is likely no evidence of fraud using the Benford Analysis method. The frequencies of each leading digit in the financial transaction dataset are too similar to the optimal leading digit frequencies (as proposed by Benford) to warrant further investigation. 
```{r HWQ1.5}
BenfordTable <- read_csv("/Users/carstenjuliansavage/Desktop/R Working Directory/Accounting/Accounting Data Analytics/Data Files/P-Card/benford_frequencies.csv")
BenfordTable %>%
rename("Leading Digit" = "leading_digit") %>%
rename("Expected Frequency" = "expected_frequency") %>%
mutate(`Proportion %`=(`Expected Frequency`)*100)
```

```{r HWQ2}
osu14 %>%
group_by(Cardholder_First_Initial,Cardholder_Last_Name,Amount,Vendor,Transaction_Date) %>%
mutate(Count=n()) %>%
filter(Count>1) %>%
arrange(Transaction_Date,Vendor)
```

```{r HWQ3}
osu14$Amount <- as.character(osu14$Amount)
osu14 %>%
group_by(Cardholder_First_Initial,Cardholder_Last_Name,Amount,Transaction_Date) %>%
mutate(Count=n()) %>%
ungroup() %>%
group_by(Cardholder_Last_Name, Cardholder_First_Initial) %>%
summarize(TotalCount=sum(Count)) %>%
arrange(desc(TotalCount))
```

```{r HWQ4}
osu14$Amount <- as.character(osu14$Amount)
#separate(Amount, into=c("Dollars","Cents"),sep=".",convert=TRUE) #No amounts with cents exist, so this is not needed
osu14 %>% 
mutate(`Round Amount?`=ifelse(str_detect(osu14$Amount,".000$"),"Round","Not Round")) %>%
filter(`Round Amount?`=="Round") %>%
arrange(Vendor,Cardholder_Last_Name,Cardholder_First_Initial)


```

```{r HW3, Q5}
# Homework 3, Q1

# Making sure that we can merge the two tables
benford_frequencies <- read.csv(file='benford_frequencies.csv') %>% as_tibble()
benford_frequencies <- benford_frequencies %>% 
  mutate(leading_digit = as.character(leading_digit))

freq_table <- osu14 %>% 
  filter(Amount > 1.00) %>% 
  mutate(leading_digit = str_sub(Amount, 1, 1)) %>% 
  group_by(leading_digit) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  mutate(across(prop, round, 3))

freq_table %>% left_join(benford_frequencies) %>%
    rename("Leading Digit"="leading_digit") %>%
  rename("Expected Frequency"="expected_frequency") %>%
  rename("Proportion"="prop")
```

