---
title: "IntegrateCo - Homework"
author: "George Batta"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(knitr)
library(lubridate)
payroll_2015=as_tibble(read_delim("Analytics_mindset_case_studies_IntegrateCo_2015payroll.csv",col_names=TRUE,delim="|"))
payroll_2016=as_tibble(read_delim("Analytics_mindset_case_studies_IntegrateCo_2016payroll.csv",col_names=T,delim="|"))
job_codes=as_tibble(read_delim("Analytics_mindset_case_studies_IntegrateCo_jobcodes.csv",col_names=T,delim="|"))
location=as_tibble(read_delim("Analytics_mindset_case_studies_IntegrateCo_locations.csv",col_names=T,delim="|"))


#Extract job_code from 2015 data & Create Year variable
payroll_2015<- payroll_2015 %>% mutate(sub=substring(Job,3,6)) %>% 
  mutate(sub_tf=str_detect(sub,"^[a-zA-Z]")) %>% 
  mutate(job_code=ifelse(sub_tf==TRUE,substring(Job,4,6),substring(Job,3,5))) %>% 
  mutate(Year=2015) %>% select(-sub,-sub_tf)

# Extract job_code from 2015 data & create Year variable
# 14H010-> H010
# 14010 -> 010

payroll_2016<- payroll_2016 %>% mutate(sub=substring(Project,3,6)) %>% 
  mutate(sub_tf=str_detect(sub,"^[a-zA-Z]")) %>% 
  mutate(job_code=ifelse(sub_tf==TRUE,substring(Project,4,6),substring(Project,3,5))) %>% 
  mutate(Year=2016) %>% select(-sub,-sub_tf)



#Align variable names across the two datasets
payroll_2016 <- payroll_2016 %>% 
  rename(Job=Project ) %>% 
  rename(Emp_No=Employee_Number) %>% 
  rename(Reg_Hrs=Hours ) %>% 
  rename(OT_Hrs=OT_Hours ) %>% 
  rename(Fica_Medc=SocSec_Medicare) %>% 
  rename(Futa=FUTA ) %>% 
  rename(Suta=SUTA )


# Check on primary key assumptions given in case!!
# Pg. 2: "Each row represents a unique employee, 
# job code, and pay date/payroll_id combination."


payroll_2015 %>% count(First_Name,Last_Name,Payroll_ID,Job) %>% filter(n>1)


inspect <- payroll_2015 %>% filter(First_Name=="Cindy" & Last_Name=="Lunt") 


inspect<-payroll_2015 %>% filter(First_Name=="Cindy" & Last_Name=="Lunt") %>% 
  arrange(Pay_Date,Payroll_ID,Job)

# I can also inspect specific rows to make it easier to see what went wrong
payroll_2015 %>% count(Payroll_ID,First_Name,Last_Name,Job) %>% filter(n>1) 

inspect<-payroll_2015 %>% filter(First_Name=="Maryann" & Last_Name=="Cohen" & Job=="15H001B" & Payroll_ID=="27310006") 



# To align with 2016 presentation, let's sum up all pay-related variables, for the same set of other characteristics,
# using group_by (for non-wage variables) and summarize (for to-be-summed wage variables):
# Lastly, I ungroup() at the end...otherwise, I might get unexpected groupings for 
# later statistics I calculate!

payroll_2015<-payroll_2015 %>% 
  group_by(Home_Div,Work_Div,Payroll_ID,Pay_Date,First_Name,Last_Name,Job,Emp_No,Year,job_code) %>% 
  summarize(across((Reg_Hrs:Work_Comp),sum,na.rm = TRUE)) %>% 
  ungroup()

# Pithier version of above, where I use group_by(across()) to 
# exclude variables from grouping

payroll_2015<-payroll_2015 %>% 
  group_by(across(-(Reg_Hrs:Work_Comp))) %>% 
  summarize(across((Reg_Hrs:Work_Comp),sum,na.rm = TRUE)) %>% 
  ungroup()


# Check for duplicates. Yay, none!!
payroll_2015 %>% count(Payroll_ID,First_Name,Last_Name,Job) %>% filter(n>1) 



# All variables in 2016: Seems like ther are missing employee id information!
# Will just keep these in dataset, for now.
payroll_2016 %>%  count(Employee_Name,Emp_No,Job) %>% filter(n>1)


#Bind the rows together
payroll_total<-bind_rows(payroll_2015,payroll_2016)



# STUDENT EXERCISE
#Replace Gross_Wages with sum of Reg_Pay and OT_Pay if it's missing
#   i.e., is.na(Gross_WAGES)==TRUE
#Create Total_Hours variable

#But first, pull together first and last name in 2016 data with unite()

payroll_total <-payroll_total %>% 
  unite(Employee_Name1,c(First_Name,Last_Name),sep=" ") %>% 
  mutate(Employee_Name=ifelse(is.na(Employee_Name)==TRUE,Employee_Name1 ,Employee_Name)) %>% 
  mutate(Gross_Wages=ifelse(is.na(Gross_Wages),Reg_Pay+OT_Pay,Gross_Wages)) %>% 
  mutate(Total_Hours=Reg_Hrs+OT_Hrs) %>% 
  mutate(Gross_Pay_Per_Hour=Gross_Wages/Total_Hours) %>% select(-Employee_Name1)



# Merge in job_codes  
job_codes<-job_codes %>% rename(job_code=Job_Code)


payroll_total <- payroll_total   %>% left_join(job_codes)  

# This gets me the sum of NA-valued Job_Description rows
payroll_total %>% summarize(count_na=sum(is.na(Job_Description)))
# This returns for me NA-valued rows
payroll_total %>% filter(is.na(Job_Description)==TRUE)


payroll_total %>% distinct(job_code)
job_codes %>% distinct( job_code)

# Weird job_code that doesn't show up in the job_codes file. Let's filter it out.
payroll_total <- payroll_total %>% filter(job_code!="144")  

payroll_total %>% distinct(job_code)


# Homework short answers: 
# First, good to convert Pay_Date to a date variable
payroll_total <- payroll_total %>% mutate(Pay_Date= mdy(Pay_Date))

```

##Homework questions

```{r Q1}

# Question 1
payroll_total %>% filter(Year==2015  ) %>% 
  group_by(Employee_Name,Emp_No) %>% 
  summarize(across(c(OT_Hrs,OT_Pay),sum,na.rm = TRUE)) %>% 
  filter(OT_Pay>0.0) %>% 
  arrange(-OT_Pay)
```


```{r Q2}
#Question 2
payroll_total %>% 
  filter(Year==2015) %>% 
  group_by(job_code,Job_Description) %>% 
  summarize(across(c(OT_Hrs,OT_Pay),sum,na.rm = TRUE)) %>% 
  arrange(-OT_Pay,Job_Description)
```


```{r Q3}
#Question 3
payroll_total %>% filter(Year==2015 ) %>% 
  group_by(Employee_Name,Pay_Date) %>% 
  summarize(across(c(Reg_Hrs,OT_Hrs), sum,na.rm = TRUE)) %>% 
  filter(Reg_Hrs<40.0 & OT_Hrs>0.0) %>% 
  arrange(Employee_Name,Pay_Date)
```


```{r Q4}
#Question 4
payroll_total %>% 
  inner_join(location) %>% 
  group_by(Location,Year) %>% 
  summarize(across(c(Gross_Wages,Total_Hours),sum,na.rm=TRUE)) %>% 
  ungroup() %>% 
  pivot_wider(names_from=Year,values_from=(-(Location:Year))) %>% 
  mutate(Ave_Pay_15=Gross_Wages_2015/Total_Hours_2015) %>% 
  mutate(Ave_Pay_16=Gross_Wages_2016/Total_Hours_2016) %>% 
  mutate(across(c(Gross_Wages_2015:Ave_Pay_16),round,2))

```

```{r Q4 Practice}
payroll_total %>%
full_join(location) %>%
group_by(Location,Year) %>%
summarize(across(c(Gross_Wages,Total_Hours,Reg_Hrs,OT_Hrs),sum,na.rm=TRUE)) %>%
ungroup() %>%
pivot_wider(names_from=Year,values_from=c(Reg_Hrs:OT_Hrs,Gross_Wages:Total_Hours)) %>%
mutate(AveragePay2015=Gross_Wages_2015/Total_Hours_2015) %>%
mutate(AveragePay2016=Gross_Wages_2016/Total_Hours_2016) %>%
na.omit() %>%
mutate(across(c(AveragePay2015:AveragePay2016),round,2))
```


# Additional questions 
```{r Q5}
payroll_total   %>% 
  filter(Year==2016) %>% 
  mutate(sum_pay=Gross_Wages+Futa+Suta+Fica_Medc+Work_Comp) %>% 
  summarize(sum_pay_2016=sum(sum_pay,na.rm=TRUE))
```

```{r Q6}
payroll_total %>% filter(Year==2016 & is.na(Emp_No)==TRUE) %>% 
  summarize(total_na_16=n(),total_na_gross_wages=sum(Gross_Wages))  
```

**Question 7**

This was definitely the most challenging one. You don't have to do it the exact way I did it below, but I've tried to make this answer as DRY as possible. 

Note that I included one additional option in `pivot_wider()`:

`values_fill`. I found this in the documentation for `pivot_wider()` This option tells `pivot_wider()` to code missing values in the `perc` variable as equal to zero. I could do this later via another `mutate()`, but it's quicker to do it this way.

One other thing: Note that the `group_by(Employee_Name,job_code)` affects the next two `summarize()` and `mutate()` commands. All subsequent summarize and mutates will be affected by this grouping, until I `ungroup()` it or set `group_by()` again. The other thing to note is that after the `summarize()` command, the data  are grouped by `Employee_Name` only (you can see this by running everything up and and including summarize; in the output, you can see "Groups: Employee_Name"). After `summarize()`, `group_by()` will default to grouping by the first variable you put in the `group_by()`. So the ordering does matter. Why doesn't it retain grouping by both `Employee_Name` and `job_code`? Since those two variables uniquely identify each row, there's nothing you can aggregate by grouping by those two anymore!

```{r Q7}
payroll_total %>% 
  filter(Year==2016 & is.na(Employee_Name)==FALSE) %>% 
  group_by(Employee_Name,job_code) %>%
  summarize(Gross_Wages=sum(Gross_Wages,na.rm=TRUE)) %>% 
  mutate(sum_by_emp=sum(Gross_Wages,na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(perc=(Gross_Wages/sum_by_emp)*100) %>% 
  
  # These two variables clutter up the table, so I dump them.
  
  select(-c(Gross_Wages,sum_by_emp)) %>%
  
  # The arrange() here just helps put the job_codes in order from 001 to 012 after pivot_wider. It's   not absolutely necessary.
  
  arrange(job_code) %>% 
  pivot_wider(names_from=job_code,values_from=perc,values_fill=list(perc=0.0)) %>% 
  mutate(across(-c(Employee_Name), round,2)) %>% 
  arrange(Employee_Name)  
```

Only thing to note for Question 8 is the `month(Pay_Date)` part, which we haven't yet learned about. Since I transformed `Pay_Date` into a date variable, I can use date functions like `month()` to extract information from a date. 

```{r Q8}
payroll_total %>% 
  filter(Year==2015) %>% 
  mutate(month=month(Pay_Date)) %>% 
  group_by(month) %>% 
  summarize(Gross_Wages=sum(Gross_Wages,na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(Diff_from_Ave=Gross_Wages-mean(Gross_Wages))     
```

